---
title: 'Poem & Sentiment Analysis with Blogdown '
author: Tara Kenner
date: '2018-08-04'
slug: poem-with-blogdow
categories: []
tags:
  - R
  - blogdown
---
I love to read and compose poetry and this poem is one that I wrote for my grade 7 Language Arts class. Thus, I was pleasantly suprised when I came across [Yihui's blog](https://yihui.name/en/2018/06/quote-poem-blogdown/)
because I never knew that I could integrate my passion for poetry with my passion for R. Thank you so much Yihui; I appreciate it.
What I'd like to attempt is a sentiment analysis using [Julia Stilge's Text Mining with r ](https://www.tidytextmining.com/tidytext.html) so I can see the general feeling my poem gives out.


> What We Take For Granted

> My fellow gentlemen, take some advice from me  
As I wander around eternally  
Thinking of how I paid the ultimate cost  
A silent spectator to all I have lost

> Let me start at the beginning in New York city  
Where I met a young lady, clever and pretty  
We had dinner and drinks at Le Bernardin  
And danced until dawn at the downtown scene

> A week later she informed me “I’m studying for the bar”  
And laughed as I almost tripped into a car  
“It’s not very feminine”, my words left hanging  
“This is 1964” she frowned, “the world is changing”

> Months later I proposed and we moved in together  
In a luxury penthouse just off 21’st and Heather  
She bought a hi fi phono and let it blast  
100 little records of rock, rhythm and jazz

> On the weekdays I worked my 9 to 5  
Managing numbers for the Wall Street jive  
And every night I came home to our dinner table  
And we ate her roasted chicken with sugar maple  
Oh, didn’t it seem as though I had I all?  
Wealthy and charming at functions and balls  
With a beautiful fiance, never dull never aged  
When she found employment, all of that changed

> Now, I was not a misogynistic man  
Nor did I want her dressed in gloves and fans  
But we all have a place, come destiny or doom  
And a young woman’s place is not in a courtroom

> You should have heard the screaming and the yelling  
She was deaf to all reason, deaf to my tellings  
Deaf to my warnings and deaf to my pleading  
I only ever kept a lookout for her well being

> That evening I was in a terrible rage  
Drove like a madman, revved the pedal gage  
So to the loads traffic I paid no attention  
Not seeing the cars coming from all directions

> When we collided it was as if time had froze  
Then back came reality, in three sharp blows  
The sirens, the commotion, a crescendo of screams  
The smell of blood mingling with petrol gasoline  
For as long as I could remember, I had always feared dying  
But I felt no fear, not at all terrifying  
The emotion that filled me was bitter regret  
Of all of the things that her and I had said

> Mourning for the life I could have lived  
For I could take and take but never give  
Regret for not appreciating, opposed to being glad  
Regret for the best thing that I never had

>

Now for my sentiment analysis on a set of my poems that I have saved as a pdf file. Let's load the required libraries. 
```{r, error=FALSE, message=FALSE, warning=FALSE}

library(tidyverse)
library(pdftools)
library(tidytext)
library(wordcloud)
library(reshape2)
```


Now I will load my poems from the pdf file
```{r}
tara_poems<-pdf_text("../../static/data/Poems_TK.pdf")
 table_tara_poem <- data.frame(text = unlist(strsplit(tara_poems, "\n"))) %>% 
  mutate(poem = "all_my_poems", line = row_number(), text = gsub("\n", "", text)) 

#head(table_tara_poem)
tara_poem_text<- table_tara_poem %>% 
    as_tibble() %>% 
   unnest_tokens(word, text) %>% 
   anti_join(stop_words)

tara_poem_text
```

```{r, error=FALSE, message=FALSE, warning=FALSE}
tara_poem_text %>% 
  inner_join(get_sentiments("bing")) %>% 
  count(word, sentiment, sort = TRUE) %>% 
  acast(word~sentiment, value.var = "n", fill = 0) %>% 
  comparison.cloud(colors = c("gray20", "gray80"), max.words = 100)

```




