---
title: Functions & Graphs
author: Tara Kenner
date: '2018-08-06'
slug: functions-graphs
categories:
  - math
  - data
  - Algebra
tags:
  - Equations
  - ggplot
---

So over the summer, I took some Algebra/Advanced Algebra classes at Saturday Academy at PSU.  We covered Exponents, Logarithms, matrices, polynomials, quadratics, solving systems of equations w/different methods, Factorisation, Graphing of functions,compound interest......

My favorite: Matrices, and graphs, so I'll start out with writing functions and plotting them in R.

Linear equation in slope intercept form:

```{r}
library(tidyverse)

func_parabola <- function(x) {
  x^2
}

func_parabola = function(x){
  x^2
}

plot(func_parabola, 2,-2)
domain_data <- tibble(x = c(2,-2))

plot(func_parabola(c(2,-2)))

domain_data %>% ggplot(aes(x = x)) + stat_function(fun = func_parabola)

```

```{r}

domain_data_1 <- tibble(x = c(2,4))
sqroot_function <- function(x){
  if(x>=0) {
    sqrt(x)
  }
}

domain_data_1 %>% ggplot(aes(x =x)) + stat_function(fun = sqroot_function)


```

