---
title: Forces
author: Tara Kenner
date: '2018-02-25'
slug: forces
categories:
  - Physics
  - r
  - math
tags:
  - blogdown
  - ggplot
  - highcharter
  - Math
  - Equations
---

In my Physics class, we are learning about Force and Newton's Laws. 

I've learned quite a bit about the difference between `speed`, `velocity`, `acceleration`.  `Speed` is distance/time (measured in m/sec: how far one travels a meter in 1 sec or whatever unit of distance and time specified specified.  It is a scalar), `Velocity` is speed with direction (so it is a vector; has both direction and magnitude), and `acceleration` is the rate of change of velocity (it is a vector, and measured in m/sec^2.  So acceleration is the change in m/sec/sec). 

I've also learned the difference between `mass` and `weight`; `mass` is the total substance in a body (measured in KG/lb, and it is a scalar; only has magnitude), while `weight` is a force which is measured in Newtons. It is how hard gravity pulls on an object.  

I will create a function that calculates Force in Newtons given Mass and Acceleration. This is Newton's second law of motion `F = m*a` where, F is the force, m is the mass and a is the acceleration.

```{r}

#The function
ForceEquation <- function(mass, acceleration) {
  
  F = mass*acceleration
  return(F)
}

#Call the function ForceEquation with mass 6 Kg, and acceleration @ 22 m/sec^2
ForceEquation(6, 22)
```

Using the same second law of motion equation, I can calculate the mass M.

```{r}

MassEquation <- function(force, gravity_a) {
  
 M = force/gravity_a 
  return(M)
}

#Call the MassEquation Function with Force 196 Newtons, and gravitational acceleration at 9.8m/sec^2
MassEquation(196, 9.8)
```






